---
title: "lab_02"
author: "derek willis"
date: "2023-09-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most?

**A1. Atmospheric and Oceanic Science offered the most classes with Climate in the title and at least one seat. Atmospheric and Oceanic Science offered 68, and there are 26 department which fit this criteria.**

```{r}
umd_courses |>
  filter(str_detect(title, "Climate") & seats > 0) |>
  group_by(department) |>
  summarize(classes = n()) |>
  arrange(desc(classes))

```

**Q2.** Of the departments listed in A2, which one is the most interesting to you, and why?

**A2. Art History & Archaeology is the most interesting to me, as there isn't an explicit reason that I think Art History would mention climate.**

**Q3** Which departments have the most courses in the Fall 2023 term? And which departments in that term have at least one syllabus posted? How are the top five departments different in each result? Where is Journalism in those results?

**A3. Business and Management had the most courses in Fall 2023, with 82 courses. 125 different departments have a syllabus posted for one of their classes. Business was definitely an outline, having double the amount of courses that the 4th and 5th departments did. I also found the departments to be fairly difference, with Economics and the School of Music being side by side. Journalism is the 38th department, tied with a couple of other departments.**

```{r}
umd_courses |>
  group_by(department) |>
  filter(term == 202308 & syllabus_count > 0) |>
  summarize(courses = n()) |>
  arrange(desc(courses)) 

```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as the same dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.

**A4. The Wes Moore committee spent the most in October of 2022. The spending generally increased from September 2022 to July of 2023, where it evened out a bit until hitting an all-time high in October.**

```{r}
maryland_expenses |>
  mutate(expenditure_date = mdy(expenditure_date)) |>
  mutate(month = floor_date(expenditure_date, "month")) |>
  filter(committee_name == 'Moore  Wes For Maryland') |>
  group_by(month) |>
  summarise(total_amount = sum(amount)) |>
  arrange(desc(total_amount))
  
```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for the others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe).

**A5: The results added up the amount of in-state contributions and out of state contributions, showing that the vast majority came from in-state donors.**

```{r}
maryland_expenses |>
  mutate(
    spending_location = case_when(
      str_detect(address, " Maryland ") ~ "In-State",
      str_detect(address, " MD ") ~ "In-State",
      str_detect(address, " California ") ~ "Out of State",
      str_detect(address, " Washington ") ~ "Out of State",
      str_detect(address, " Louisiana ") ~ "Out of State",
      str_detect(address, " Florida ") ~ "Out of State",
      str_detect(address, " North Carolina ") ~ "Out of State",
      str_detect(address, " Massachusetts ") ~ "Out of State",
      str_detect(address, " West Virginia ") ~ "Out of State",
      str_detect(address, " Virginia ") ~ "Out of State",
      .default = NA
    ) 
  )|>
  
    group_by (spending_location)|>
    summarize (total_amount=sum(amount))
  
```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.

**A6. How many seats were open in each term of 2023? Answer: 221,036 in Fall semester, 208,465 in Spring semester, and 42,791 in Summer semester.**

```{r}
courses_2023 |>
  mutate(
    term_total = case_when(
      str_detect(term, "202305") ~ "Summer 2023",
      str_detect(term, "202301") ~ "Spring 2023",
      str_detect(term, "202308") ~ "Fall 2023",
      .default = NA
    ) 
  )|>
  
group_by(term_total) |>
  summarize(total = sum(seats)) |>
  arrange(desc(total))
```
