---
title: "pre_lab_06.Rmd"
author: "derek willis"
date: "2023-03-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Points to hit

1.  Review of previous lab questions/problems.
2.  Demonstration of combining and merging

## Chapter 23

### Task 1: Load libraries and settings

**Task** Run the following code in the gray-colored codeblock below to load the tidyverse library and turn off scientific notation.

```{r}
# turn off sci notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)
```

### Task 2: Load data

Let's say that we have Maryland county voter registration data from four different elections in three different files. They have a similar layout and the same number of counties (plus Baltimore City).

**Task** Load the four CSV population estimate files. Do they have the same column structure?

**Answer: They do have the same column structure, except in 2022 when the last column is called "Total" instead of "TOTAL"**

```{r}
county_voters_2016 <- read_csv("data/county_voters_2016.csv") |>
  clean_names()
county_voters_2018 <- read_csv("data/county_voters_2018.csv") |>
  clean_names()
county_voters_2020 <- read_csv("data/county_voters_2020.csv") |>
  clean_names()
county_voters_2022 <- read_csv("data/county_voters_2022.csv") |>
  clean_names()
```

### Task 3: Using bind_rows

**Task** Combine the four files into a single data frame using the bind_rows function. Add a description of what this code does to your reference notebook.

```{r}
# bind_rows with list
county_voters_combined <- bind_rows(list(county_voters_2016, county_voters_2018, county_voters_2020, county_voters_2022))
View(county_voters_combined)
```

### Task 4: Asking a question

**Task** Run the following code to filter on Frederick County's registration data and look at the results. What stands out to you about the changes in party registration over time?

**Answer: Overall registration has risen over time, swapping from primarily Republican to primarily Democrat.**

```{r}
county_voters_combined |> filter(county == 'Frederick')
```

## Joining data

### Task 5: Loading population data

**Task** Load the Maryland 2020 county population data

```{r}
maryland_population <- read_csv('data/maryland_population_2020.csv') |>
  clean_names()
view(maryland_population)
```

### Task 6: Join population and voter data

**Task** Run the following code to join Maryland population data and the 2022 voter registration data together using the common column county as the key. How many rows are there now? How many *should* there be?

**Answer: There are 24 rows, which is the correct amount.**

```{r}
# with nrow included to show row total
maryland_population |> left_join(county_voters_2020, join_by("county")) |> nrow()
```

### Task 7: Looking at Joined Data

**Task** Examine the combined data

```{r}
# without nrow
maryland_population |> left_join(county_voters_2022, join_by("county"))
```

### Task 8: Assign Joined Data to New Dataframe

By default, `dplyr` will do a "natural" join, where it'll match all the matching columns in both tables. So if we take out the `join_by()`, it'll use all the common columns between the tables. That may not be right in every instance but let's try it. If it works, we should get 24 rows.

**Task** Create a new dataframe with population and voter data joined together. Add a description of what this code does to your reference notebook.

```{r}
maryland_population_with_voters <- maryland_population |> left_join(county_voters_2022)

# It used the same column as before to do the joining.
```

### Task 9: Add the Demographic Data

**Task** Load the Maryland demographic data and examine it. What kind of information does it contain? **Answer: It has the number of population over 18 and the racial demographics of each county.**

```{r}
maryland_demographics <- read_csv('data/maryland_demographics.csv') |>
  clean_names()
view(maryland_demographics)
```

### Task 10: Join demographic data to combined voting/population data

**Task** Join the demographic data to the our combined voting/population data. To do this, we need to specify the join columns since the names aren't the same in the two dataframes.

```{r}
maryland_population_with_voters_and_demographics <- maryland_population_with_voters |> left_join(maryland_demographics, join_by(county == name))
```

### Task 11: Asking Demographic Questions

**Task** Using mutate, let's find the county with the highest 18+ Black population as a percentage of all population 18 and over and also the percentage of Democrats in that county. Put your answers below. **Answer: Prince George's Counted has the highest percentage of Black people over 18, with a percentage of 77% democrats.**

```{r}
maryland_population_with_voters_and_demographics |>
  mutate(pct_black_18_plus = (pop_black/pop_18_over)*100, pct_dems = (dem/total)*100) |> 
  arrange(desc(pct_black_18_plus)) |> 
  select(county, pct_black_18_plus, pct_dems)
```

### Task 12: Asking Demographic Questions

**Task** Now let's find the county with the highest 18+ Asian population as a percentage of all population 18 and over and also the percentage of Democrats in that county. How do these results compare to the previous task? Put your answers below.

**Answer: This shows that Howard County has the highest 18+ Asian population, with 53% democrats. Having a 20% Asian population is much lower than the 61% Black population showed in the previous task. They also have a much smaller democratic percentage.**

```{r}
maryland_population_with_voters_and_demographics |>
  mutate(pct_asian_18_plus = (pop_asian/pop_18_over)*100, pct_dems = (dem/total)*100) |> 
  arrange(desc(pct_asian_18_plus)) |> 
  select(county, pct_asian_18_plus, pct_dems)
```

### Task 13: Fixing Join Problems

Sometimes joins look like they should work but don't. Often this is due to the two columns you're joining on having different data types: joining a <chr> column to a <dbl> column, for example. Let's walk through an example of that using some demographic data by zip code.

**Task** Run the following code to load the Zip Code Tabulation Area data for Maryland. What's the datatype of the ZCTA5N column?

**Answer: ZCTA5N has a data type of dbl.**

```{r}
maryland_zcta <- read_csv('data/maryland_zcta.csv')
glimpse(maryland_zcta)
```

The column we're interested in, ZCTA5N, is a <dbl> column - it's a number. We want it to be a <chr> column - text.

**Task** Run the following code to change the datatype of ZCTA5N from numeric to character. What's the datatype of ZCTA5N?

**Answer: It is now a character datatype.**

```{r}
maryland_zcta <- maryland_zcta |> mutate(ZCTA5N = as.character(ZCTA5N))
glimpse(maryland_zcta)
```

Now we can join this dataframe to other zip code data where the zip code column is text, not numbers.
