---
title: "Reverse Engineering Project"
author: "Kenan Grier, Marissa Yelenik, Matthew Kiras, Nicholas Wolcott"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, Thousands of kids are missing from school. Where did they go? (<https://projects.apnews.com/features/2023/missing-children/index.html>)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

population_by_age <- read.csv('data/population-by-age.csv') |>
  clean_names()
private_enrollment_by_grade <- read.csv('data/private-enrollment-by-grade.csv') |>
  clean_names()
public_enrollment_by_grade <- read.csv('data/public-enrollment-by-grade.csv') |>
  clean_names() |>
  
missing_kids_analysis <- read.csv('data/Missing-Kids-Analysis.csv') |>
  clean_names()
school_aged_population_changes <- read.csv('data/School-Aged-Population-Changes.csv') |>
  clean_names()
enrollment_changes <- read.csv('data/Enrollment-Changes.csv') |>
  clean_names()



# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: "Those states saw private-school enrollment grow by over 100,000 students" from 2019-2020 to 2021-2022
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
private_enrollment_by_grade |>
  group_by(state) |>
  summarise(
    amt_change = 
    ) |>
  mutate(
    
  )
  arrange(desc(amt_change))

  
# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
