---
title: "Reverse Engineering Project"
author: "Kenan Grier, Marissa Yelenik, Matthew Kiras, Nicholas Wolcott"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, Thousands of kids are missing from school. Where did they go? (https://projects.apnews.com/features/2023/missing-children/index.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)


```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

population_by_age <- read.csv('data/population-by-age.csv') |>
  clean_names()

private_enrollment_by_grade <- read.csv('data/private-enrollment-by-grade.csv') |>
  clean_names()

public_enrollment_by_grade <- read.csv('data/public-enrollment-by-grade.csv') |>
  clean_names()

missing_kids_analysis <- read.csv('data/Missing-Kids-Analysis.csv') |>
  clean_names()

school_aged_population_changes <- read.csv('data/School-Aged-Population-Changes.csv') |>
  clean_names()

enrollment_changes <- read.csv('data/Enrollment-Changes.csv') |>
  clean_names()


# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: "Overall, public school enrollment fell by over 700,000 students between the 2019-2020 and 2021-2022 school years in the 21 states plus Washington, D.C., that provided the necessary data."
* **Analysis summary**: We utilized data from the data set 'enrollment_changes' and compared it to the data set 'missing_kids_analysis'. We were able to find the number of children who left public schools between the 2019-2020 and 2021-2022 school years, and it matched the number featured in the story.

```{r}
# Put code to reverse engineer sentence here

enrollment_changes |>
  group_by(enrollment_type) |>
  filter(enrollment_type == 'Public school enrollment') |>
  filter(state_abbrev == 'CA' | state_abbrev == 'NY' | state_abbrev == 'LA' | state_abbrev == 'NC' | state_abbrev == 'WA' | state_abbrev == 'CO' | state_abbrev == 'GA' | state_abbrev == 'MN' | state_abbrev == 'WI' | state_abbrev == 'NH' | state_abbrev == 'MA' | state_abbrev == 'MT' | state_abbrev == 'DE' | state_abbrev == 'ND' | state_abbrev == 'DC' | state_abbrev == 'SD' | state_abbrev == 'SC' | state_abbrev == 'NE' | state_abbrev == 'RI' | state_abbrev == 'TN' | state_abbrev == 'PA' | state_abbrev == 'FL') |>
  mutate(school_year_2019_20_21states = sum(school_year_2019_20)) |>   mutate(school_year_2020_2021_21states = sum(school_year_2020_2021)) |>
  mutate(school_year_2021_22_21states = sum(school_year_2021_22)) |>
  mutate(public_school_change_21states = school_year_2021_22_21states-school_year_2019_20_21states) 
  

# Display results of code below this codeblock

```

### Sentence 2

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here



# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-